---
title: "reso"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reso}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
devtools::load_all()
library(reso)

```

# reso: An esoteric R-based programming style

### Introduction 

Have you ever been frustrated at R, maybe after reading too many rude
Stackoverflow or R-help threads? (me too) Well, what if we took R and instead
of helping making it accessible and fun, made all of those esoteric,
argon-y, incredibly annoying, and unreadable parts much worse?

Enter `reso`, a library and R-based programming style. Only the
following symbols from R are allowed:

<p style="text-align:center">
(), [], !, 0, +, and "," (the comma).
</p>

Along with four functions to make `reso` "go." They are really
just simplified and narrower versions of existing R utilities:

- s(): An alias for substr(string, i). Used to get a letter from a string

- p(): An alias for paste(..., collapse = ""), to concatenate strings

- x(): Interpret and execute a string as an R expession

- d(): an alias for [deparse()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/deparse)

And that's it! Not even spaces, line breaks, or tabs are allowed! Happy programming :)

### How it works

Just kidding, there's a few more things that should be said. 

This style is able to work because R is able to interpret 
strings as expressions, and through hacky (some might say c l e v e r) 
tricks you can grab those strings from native R objects that are available to you. 

The general pattern using this is as follows:

1. Find a hacky way to coerce values to built-in objects, like `FALSE` or `logical(0)`
2. Interpret that value as a string with `d()`, so that `FALSE` becomes `"FALSE"` and so on.
3. Pull a letter of that string out with `s()`
4. Repeat and combine with `p()` until we can make a function name
5. Evaluate the string as an R expression with `x()`. 
6. If that worked, you have a new function available to you! 
Repeat ad infinitum until your code is completely nonsensical.

### Caveats 

There are some things that I'm not sure if the style can do as of now. 
Probably the biggest issue is that a programmer can't save variables to the namespace.
The `reso_validate()` function is super hacky, such that it just marks as valid those
strings or file expressions which have only `reso` characters. This means the 
assignment to new characters is not allowed. It'd also break because the `d()` function operates 
in the parent frame instead of in the global environment with no escape hatch. The parent 
frame would most often be the context inside of `x()` and would be "forgotten" after execution,
meaning that variables can't really be saved.

### A demonstration

**Goal: Generate the output `[1] 0 1`** 

Or in other words, the output we would receive if we were typing `c(0, 1)` in vanilla R. 

Let's start by getting just the letter "c" as a string, which we can do by:
1. Produce `character(0)` by generating an empty string with R's 
`paste` or reso's `p`, then get the 0th element of that result. 
2. Convert the value `character(0)` into character, e.g. 
`"character(0)"`. This can be done with R's `deparse` or reso's `d`. 
3. Get the first letter of "character(0)", 
via R's `substr` or reso's `s`. 

Those steps as code: 

```{r}
p()[0]
d(0[0])
s(d(p()[0]))
```

Next, we'd want to turn this string into the R expression `c(0, 1)`. To do this, we can: 

1. Force R to interpret "c" as the native function `c`, by using `parse` and then `eval` in R, or using the `x` function^[for "execute", or "expression." Your choice :)] in reso. 
2. Generate 1 by coercing 0 to `TRUE` with `!0`, and combining `TRUE` with `numeric(0)` to force `TRUE` to be numeric (or 1). This requires our new version of `c()`.
3. Combine all of our above steps with a set of parens and a 0!

```{r}
x(s(d(p()[0])))
x(s(d(p()[0])))(!0,0[0])
x(s(d(p()[0])))(0,x(s(d(p()[0])))(!0,0[0]))
```

Yay! 

### Expanding to almost the entire R universe

#### Part 1: The alphabet

This could have been a bit of a puzzle to figure out,
and potentially impossible.
R has a fair number of native values, but many letters are just
hard to figure out or produce via type coercion. 

Luckily, R provides an object that makes this almost trivial: the `letters` array! 

To get it, we need: 

1. "l", from `logical(0)`
2. Two "e"s, from `integer(0)`
3. Two "t"s, from `integer(0)`
4. One "r", from `integer(0)`
5. One "s", from the eleventh letter of the deparse function expression 
for `s()`.^[If you can think of a better way, let me know :')] 

The code representation of this is below. `c()` and numbers are
used instead of their `reso` equivalents for readability.^[LOL]

```{r}
s(d(!0[0])) # the first letter of "logical(0)", or "l"
c() + !0 # makes integer(0) by adding TRUE and NULL
s(d(c() + !0, 4)) # turns integer(0) -> "integer(0)" -> "e"
s(d(c() + !0), 4) #"t"
s(d(c() + !0), 7) #"r"
s(d(s)[1], 11) #"s"

#combining
x(p(
  s(d(!0[0])),
  s(d(c() + !0), 4),
  s(d(c() + !0), 3),
  s(d(c() +  !0), 3),
  s(d(c() + !0), 4),
  s(d(c() + !0), 7),
  s(d(s)[1], 11)
))
```

We can subset letters from here in order to form arbitrary words, and thus nearly all R commands. 

### Part 2: Non-alphabetical characters

The last trick in getting this to be an almost-all-of-R style
is in using the `builtins` vector. `builtins` contains the names 
of all built-in functions in R, many of which include non-alphabetical
operators like +, -, (, :, and so on. You could subset the builtins array
 for these, which would then unlock the door to almost all of R. 

But hey, we get a lot from just twelve characters!

### Some notes about motivation and learning

This is not meant to be a useful or practical programming tool. It is meant to show the opposite, that being esoteric is useless and using simpler, abstracted logic can ^[of course, it was also just a fun/silly little side project for myself.]

Compared to languages like JavaScript, R stands out in being more accessible and friendly. JavaScript programmers attempt to be as concise as possible, preferring arrow functions to jam as much code as possible into a single line. Logical operators like '&&' and '||' are used via a side effect as if statements themselves instead of just inside them. 

R instead views being verbose as a feature. The base R libraries export thousands of functions to handle many general cases and provide a higher level of abstraction. Pipes in R are encouraged in order to demystify code and reduce mistakes, and tools like the R linter even encourage placing every statement after a pipe on a new line to encourage even more readability. 

Of course, R still has to go a long way and JavaScript has had its own readability efforts for many years. 



